---
title: "Regression Homework - homework 2"
author: "Alisha Gumber"
date: "Fall 2019"
output:
  html_document:
    df_print: paged
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T)
```

# Lab Section

In this lab, we will go over regression. We will be using the caret package in R. https://topepo.github.io/caret/train-models-by-tag.html

# Perfomance Metrics 

## Residual 

Deviation of the observed value to the estimated value (sample mean)
$$residual=y_i - \hat{y_i}$$
where $\hat{y_i}$ is the estimated value

## Mean Squared Error (MSE)

$$MSE=\frac{1}{n}\sum_{i=1}^{n}(y_i - \hat{y_i})^2$$

## Root Mean Squared Error (RMSE)
Same units as original data.

$$RMSE=\sqrt{MSE}$$

## L2 regularization : Ridge regression. Regularize by adding the sum of the coefficients, squared, to the function. 

$$Ridge Regression=\sum_{i=1}^{n}(y_i - w_0 - \sum_{j=1}^{p}w_jx_{ij})^2 + \lambda\sum_{j=1}^p(w_j)^2$$

## L1 regularization : Lasso Regression. Regularize by adding the sum of the absolute value of the coefficients to the model. Coefficient estimates may be pushed to zero -- Lasso can perform variable selection

$$Lasso Regression=\sum_{i=1}^{n}(y_i - w_0 - \sum_{j=1}^{p}w_jx_{ij})^2 + \lambda\sum_{j=1}^p|w_j|$$


\newpage

### The broad steps of Machine learning in R. 

1. Split the data into training and test. Set test aside. 

2. Fit a good model to the training data. 

3. See how your model did on the training data.

4. Test how your model performs on the test data. 

# Regression

```{r, include=FALSE}
library(caret)
library(MASS)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(glmnet)


#Mauna Loa CO2 concentrations
data(airquality)
```

```{r}
#View(airquality)
```

1. Split data into training and test set (75% in train set, 25% in test set)

```{r}
## 75% of the sample size

sample_size <- floor(0.75 * nrow(airquality))

## set the seed to make partition reproducible
set.seed(115)
training <- sample(seq_len(nrow(airquality)), size = sample_size)

train_regression = airquality[training, ]
test_regression = airquality[-training, ]
head(train_regression) # 114 observations
head(test_regression) # 39 observatons
```


### Linear Regression

* Assumes a linear relationship. 
* Independent variables should not be correlated (no mulitcollinearity)
* The number of observations should be greater than the number of independent variables.


$$RSS=\sum(y_i - \hat{y_i})^2$$
We will predict the response of the Temperature based on Wind. 

This is the data we will fit a linear model to. 
```{r}
# response of temp based on wind means y is dependent on x so x = wind (independent variable) and y = temp (dependent variable)
train_plot <- ggplot(data = train_regression) +
   geom_point(aes(x=Wind, y=Temp)) + 
   theme_bw()
train_plot
```

2. Create and fit a linear model to predict Temperature from Wind using the training set

```{r}
#help(train)

linear_regression <- train(Temp ~ Wind, data=train_regression, method = "lm")
summary(linear_regression)
```


3. Vizualize how your model performed on the train data by plotting the regression line on top of the train data points. 
```{r}
# use intercept and slope values from summary of the linear regression model
# y = 89.9657 + -1.2198x 
train_plot + geom_abline(intercept = 89.9657, slope = -1.2198, col="red") 
# + geom_smooth(method = 'lm', color=)
```


4. Explore how the model performs on the test data. For Linear Regression:

* The residuals should be close to zero.
* There should be equal variance around the regression line (homoscedasticity).
* Residuals should be normally distributed.
* Independent variables and residuals should not be correlated.

4 a) See how the model performs on the test data
```{r}
#help(predict)
linear_predict <- predict(linear_regression, test_regression)
linear_predict
```

4 b) Look at the residuals. Are they close to zero?
```{r}
#look at the median residual value. Close to zero is best
summary(linear_regression)
```
**The median redidual is 1.756, which is pretty close to zero, but there is a relatively large range of residuals.**

```{r}
# on my own: calculating error
error <- linear_predict - test_regression$Temp
RMSE = sqrt(mean(error^2))
RMSE
```


4 c) Plot predicted temperature vs observed temperature. A strong model should show a strong correlation
```{r}
plot(test_regression$Temp, linear_predict, main="Plot of predicted temp vs. observed temp")

```


```{r}
cor.test(test_regression$Temp, linear_predict)
```


4 d) Visualize the predicted values in relation to the real data points. Look for homoscedasticity
```{r}
# Extract coefficients from the model
coef1 <- linear_regression$finalModel$coefficients[1]
coef2 <- linear_regression$finalModel$coefficients[2]

# plot the regression line on the predicted values
plot(x=test_regression$Wind, y=linear_predict, pch=1, main="Regression line on predicted values") 
abline(coef1, coef2, col="red")

# plot the original test values
ggplot(test_regression) +
  geom_point(aes(x=test_regression$Wind, y=linear_predict, col="Predicted")) +
  geom_point(aes(x= Wind, y= Temp, col="Observed")) +
  geom_segment(aes(x=Wind, y=Temp, xend=Wind, yend=linear_predict))
```


4 e) Residuals should be normally distributed. Plot the density of the residuals
```{r}
residuals_lin <- residuals(linear_regression)
head(residuals_lin)

ggplot(x=residuals_lin) +
  geom_density(aes(residuals_lin))
```


4 f) Independent variables and residuals should not be correlated
```{r}
cor.test(train_regression$Wind, resid(linear_regression))
# Correlation test shows indepdent variables and residuals are not correlated
```

### Linear Regression with Regularization

5. Create a linear model using L1 or L2 regularization to predict Temperature from Wind and Month variables. Plot your predicted values and the real Y values on the same plot. 

```{r}
# cross validation with sample of 10 observations
control =  trainControl(method = "cv", 10)

# train model with ridge regression (L2)
ridge_regression <- train(Temp ~ Wind + Month, data = train_regression,
                          method = 'ridge', trControl= control) 
ridge_regression

# examine residuals
ridge_test_predict <- predict(ridge_regression, newdata = test_regression)

#plot the predicted values vs the observed values
plot_ridge_test_pred <- data.frame(Temp_test_pred = ridge_test_predict, 
                                   Observed_Temp = test_regression$Temp)
ggplot(data = plot_ridge_test_pred) +
  geom_point(aes(x=Observed_Temp, y = Temp_test_pred)) + 
  ggtitle("True Temp Value vs Predicted Temp Value with Ridge Regression") +
  theme_bw()

#median residual value should be close to zero
median(resid(ridge_regression))
```

